---
title: "2015-04-18-Modify-Repo"
author: "Julin Maloof"
date: "April 18, 2015"
output: html_document
---

This is a file to demonstrate how to add folders and/or files to a set of github repositories and then to push those changes.

```{r init}
repos.dir <- "~/git/r_club_members" # root directory for the repositories
# repos.dir <- "~/Work/Code/BIS180L_Grading" # for Kristen
library(github)
library(plyr)
ctx <- create.github.context(access_token = Sys.getenv("GITHUB_TOKEN"),verbose=F)
```

```{r get_repos}
repos <- get.organization.repositories(org="UCD-pbio-rclub",ctx=ctx)
repo.names <- sapply(repos$content,function(x) x$name)
```

reduces repository list to those that match our keyword
```{r filter_repos}
keyword <- "Rclub-rethinking_"
repos$content <- repos$content[grepl(keyword,repo.names)]
repo.names <- sapply(repos$content,function(x) x$name)
```

pull any changes into the current repositories
```{r pull_repos}
for(repo in repo.names) {
  repo.path <- file.path(repos.dir,repo)
  error <- system(paste("cd",repo.path,"; pwd; git pull")) #error will be 0 if success, 1 if failure
  if(error) warning(paste("Error pulling into", repo.path))
  }
```

Add a directory (if necessary)
```{r add_dir}
new.dir <- "Assignment_Chapter_05"
for(repo in repo.names) {
  new.dir.path <- file.path(repos.dir,repo,new.dir)
  if(! file.exists(new.dir.path)) dir.create(new.dir.path)
  
  # create a .gitignore file so that the directory can be added to the repository.  
  #The "cat" command in this form creates an empty file.
  if(! file.exists(file.path(new.dir.path,".gitignore"))) cat(file=file.path(new.dir.path,".gitignore"))
  }
```

Delete a directory (if necessary)
```{r add_dir, eval=FALSE}
del.dir <- "Assignment_Chapter_04"
for(repo in repo.names) {
  del.dir.path <- file.path(repos.dir,repo,del.dir)
  if(dir.exists(del.dir.path)) system(paste("rm -r ",del.dir.path))
  }
```


Commit and push empty directory
```{r push_repos}
for(repo in repo.names) {
  repo.path <- file.path(repos.dir,repo,new.dir)
  if(! file.exists(file.path(new.dir.path,".gitignore"))) cat(file=file.path(new.dir.path,".gitignore"))
  error <- system(paste("cd",repo.path,
                        "; git add",file.path(repo.path,"*"),
                        "; git commit -m 'added folder",new.dir,
                        "'; git push")) #error will be 0 if success, 1 if failure
  if(error) stop(paste("Error pushing into",repo.path))
  }
```


Copy file(s) to a repo directory
```{r copy.files}
dest.dir <- "Assignment_Chapter_05/"
source.file.dir <- "~/git/r_club/_assignment_templates/"
source.files <- dir(source.file.dir,full.names=T,pattern = "Chapter-05-part2-assignment.Rmd")
if (length(source.files)==0) stop(paste("Error",source.file.dir,"empty"))
source.files.string <- paste(source.files,collapse=" ")
for(repo in repo.names) {
  repo.path <- file.path(repos.dir,repo,dest.dir)
  error <- system(paste("cp -n",source.files.string,repo.path)) #error will be 0 if success, 1 if failure
  if(error) stop(paste("Error copying files from",source.file.dir,"to",repo.path))
  }
```

commit and push changes
```{r push_repos}
for(repo in repo.names) {
  repo.path <- file.path(repos.dir,repo,dest.dir)
  source.files.base.string <- paste(basename(source.files),collapse=" ")
  error <- system(paste("cd",repo.path,
                        "; pwd; git add",source.files.base.string,
                        "; git commit -m 'added",source.files.base.string,
                        "'; git push")) #error will be 0 if success, 1 if failure
  if(error) stop(paste("Error pushing into",repo.path))
  }
```

commit all and push changes
```{r push_repos}
for(repo in repo.names) {
  repo.path <- file.path(repos.dir,repo)
  error <- system(paste("cd",repo.path,
                        "; pwd; git commit -am 'JNM Added first assignment directory'; git push")) #error will be 0 if success, 1 if failure
  if(error) stop(paste("Error pushing into",repo.path))
  }
```

get repository commits
```{r get commits}
#get the commits
commits <- list()
for(repo in repo.names) {
commits[[repo]] <- get.repository.commits("UCD-pbio-rclub",repo,ctx)
  }

#check to see if there were any errors
if (all(sapply(commits,function(x) (x$ok)))) {
  print("All commits downloaded successfully")
  } else {
    names(sapply(commits,function(x) (x$ok)))[!sapply(commits,function(x) (x$ok))]
          }

#commits is a list with a length equal to the number of repos
#go through each repo
commit.results <- sapply(commits,function(repo.commits) {
  #for each repo get the commits and extract info
  t(sapply(repo.commits$content, function(commit) {
    c(name=commit$commit$committer$name,
      msg=commit$commit$message,
      time=strftime(as.POSIXct(
        #this is ugly but it manages to convert to out timezone
        sub("T"," ",commit$commit$committer$date),"GMT"),"")
      )
    }
    ))
  }
  )

commit.results$BIS180L_Assignments_Andy.Peng

#or maybe you prefer it in one table:

commit.results.table <- ldply(commit.results)
head(commit.results.table)
```
