---
title: "R Github"
output: html_document
---

# Github and R 

I need programatic interaction with github so that I can set up user repositories, etc

## References

[GitHub API](https://developer.github.com/v3/)  
[R GitHub bindings](https://github.com/cscheid/rgithub)

## setting up

Only needs to be done once (from R).  

```{r install}
library(devtools) #Library that allows installation of libraries from GitHub
install_github("jnmaloof/rgithub",ref="develop") #Julin's fork that fixes a big in getting group repositories. Also allows download of up to 100 repos at a time (instead of default 30)

#install_github("cscheid/rgithub") #original repository
```
  
  
  
## Trying it
```{r downgrade httr}
install_version("httr","1.0.0")
```

First generate a token by going to https://github.com/settings/applications and then selecting "Personal Access Token > Generate New Token".  

Do NOT share the token with anyone or place it in a publically readable place.  You can follow [this recommendation](http://stat545-ubc.github.io/bit003_api-key-env-var.html) instead to not expose it in your scripts.

Initialize the context.  This effectively logs the API into your account
```{r init}
library(github)
# note: setting verbose=TRUE below can help with trouble shooting as it allows 
ctx <- create.github.context(access_token = Sys.getenv("GITHUB_TOKEN"),verbose=F)
```

List repositories
```{r list_repos,eval=FALSE}
my.repos <- get.my.repositories(ctx)
names(my.repos) #the main info is in "content"
names(my.repos$content[[1]])
unlist(lapply(my.repos$content,function(x) get("full_name",x)))
```

Create a team and add a member
```{r create_Team,eval=FALSE}
new.team <- create.team(org="UCD-pbio-rclub",content=list(name="testteam3",description="testing team3",permission="push"),ctx)
get.organization.teams("UCD-pbio-rclub",ctx)

#important: the team is referred to by its id (which is a number), not its name!!
add.member.results <- add.member.to.team(new.team$content$id,user="jnmaloof",ctx)
```

Delete a team
```{r delete_team,eval=FALSE}
delete.team(new.team$content$id,ctx)$ok
```

Create a repository
```{r create_repos,eval=FALSE}
create.repo.output <- create.repository(name="my_new_repos2",ctx=ctx)
```

## Creating repository for student assignments

For this I will

* download the list of github usernames
* for each user
  * create a team with the user name
  * add the user to the team
  * create a private repository and have the team assigned to it

Get class GitHub Usernames
```{r get_users,eval=FALSE}
library(googlesheets)
gs_rclub <- gs_title("R Club Registration WQ 2016 (Responses)")
users <- gs_read(gs_rclub)
head(users)
users$team.name <- make.names(paste(users$First.Name,users$Last.Name))
```

Get existing teams on GitHub and compare to imported spreadsheet
```{r get_existing_teams,eval=FALSE}
teams <- get.organization.teams("UCD-pbio-rclub",ctx)
current.teams <- data.frame(team.name=sapply(teams$content, function(x) x$name),
                            team.id=sapply(teams$content, function(x) x$id))

new.teams <- users[! users$team.name %in% current.teams$team.name,]
new.teams <- new.teams[duplicated(new.teams$github.username),]
```

Create new teams
```{r new_teams,eval=FALSE}
create.team.out <- apply(new.teams,1,function(team) {
  create.team(org = "UCD-pbio-rclub",
              content = list(name=team[["team.name"]],
                             description=paste("Team for",team[["First.Name"]],
                                               "to turn in Rclub hommework"),
                             permission="push")
              ,ctx)
  })
new.team.results <- as.data.frame(t(sapply(create.team.out,function(x) c(x$ok,x$content$name,x$content$id))))
names(new.team.results) <- c("ok","team.name","team.id")
new.team.results
new.teams <- merge(new.teams,new.team.results,by="team.name")
```

Add members to teams
```{r Add_members,eval=FALSE}
add.members.out <- apply(new.teams,1,function(team) {
  if (team[["ok"]]) add.member.to.team(id = team[["team.id"]],
                                  user = team[["github.username"]],
                                  ctx)
  })
all(sapply(add.members.out,function(x) x$ok))
```

Create repository for new or exisiting teams
```{r new_repositories,eval=FALSE}
repo.prefix <- "Rclub-rethinking"
repo.description <- "Repository for Rclub Statisical Rethinking"

################################################################################
## Choose one of the two below and comment the other out
#teams <- current.teams[!grepl("(Owners|Teachers)",current.teams$team.name),]
teams <- new.teams[as.logical(new.teams$ok),]
################################################################################

create.repository.out <- apply(teams,1,function(team) {
  create.organization.repository(
    org="UCD-pbio-rclub",
    name=paste(repo.prefix,team[["team.name"]],sep="_"),
    description=paste(repo.description,"from",team[["team.name"]]),
    #private=TRUE,
    team_id=team[["team.id"]],
    ctx=ctx)
  })
all(sapply(create.repository.out,function(x) x$ok))
```

Add teachers team to each repository (not strictly necessary because as owners of the organization we already have access)
```{r add_team,eval=FALSE}
add.team.out <- apply(teams,1,function(team) {
  add.repository.to.team(
    id=current.teams$team.id[current.teams$team.name=="Teachers"],
    owner="UCD-pbio-rclub",
    repo=paste(repo.prefix,team[["team.name"]],sep="_"),
    ctx)
})
all(sapply(add.team.out,function(x) x$ok))
```

To delete repositories (Not Run):
```{r delete_repositories,eval=FALSE}
BIS180Lrepos <- get.organization.repositories("UCD-pbio-rclub",ctx)
BIS180Lrepos.names <- sapply(BIS180Lrepos$content,function(x) x$name)
for (r in BIS180Lrepos.names[!grepl("gh-usernames",BIS180Lrepos.names)]) { # I want to keep the first one
  delete.repository("UCD-pbio-rclub",r,ctx)
  }
```

To delete teams (Not Run):
```{r delete_teams,eval=FALSE}
BIS180Lteams <- get.organization.teams(org = "UCD-pbio-rclub",ctx)
BIS180Lteam.names.ids <- data.frame(
  name=sapply(BIS180Lteams$content,function(x) x$name),
  id=sapply(BIS180Lteams$content,function(x) x$id))
BIS180Lteam.ids.remove <- BIS180Lteam.names.ids$id[grep(".",BIS180Lteam.names.ids$name,fixed=TRUE)] #remove the teams I want to keep from the list.
for (id in BIS180Lteam.ids.remove) {
  delete.team(id,ctx)
  }
```

