---
title: "R Github"
output: html_document
---

# Github and R 

I need programatic interaction with github so that I can set up user repositories, etc

## References

[GitHub API](https://developer.github.com/v3/)  

[gh package](https://github.com/r-pkgs/gh)

## setting up

Only needs to be done once (from R).  

```{r install}
library(devtools) #Library that allows installation of libraries from GitHub
install_github("r-pkgs/gh")

```
## Trying it

First generate a token by going to https://github.com/settings/applications and then selecting "Personal Access Token > Generate New Token".  

Do NOT share the token with anyone or place it in a publically readable place.  You can follow [this recommendation](http://happygitwithr.com/api-tokens.html) instead to not expose it in your scripts.

```{r init}
library(gh)
```

List your own repositories
```{r list_repos,eval=FALSE}
my.repos<- gh("/user/repos", type = "all")
vapply(my_repos, "[[", "", "name") 
sapply(my.repos,function(x) x[["name"]]) #more familiar version, same result
sapply(my.repos,function(x) x[["full_name"]]) #get longer name
```

Create a team and add a member
```{r create_Team,eval=FALSE}
team.create.out <- gh("POST /orgs/:org/teams", org="UCDBIS180L-17",name="testteam3",description="testing team3")

teams <- gh("/orgs/:org/teams", org="UCDBIS180L-17")

sapply(teams,function(x) x[["name"]])

team.id <- sapply(teams,function(x) x[["id"]])

#important: the team is referred to by its id (which is a number), not its name!!
add.member.results <- gh("PUT /teams/:id/memberships/:username",id=team.id,username="jnmaloof")
```

Delete a team
```{r delete_team,eval=FALSE}
gh("DELETE /teams/:id", id =team.id)
```

Create a repository for the authenticated user
```{r create_repos,eval=FALSE}
create.out <- gh("POST /user/repos", name="test repo132")
```

Create a repository for an organization
```{r}
out <- gh("POST /orgs/:org/repos", org="UCDBIS180L-17", name="test2")
```

Add a team to a repository

```{r}
out <- gh("PUT /teams/:id/repos/:org/:repo", id=team.id, org="UCDBIS180L-17", repo="test2", permission="push")
```



## Creating repository for student assignments

For this I will

* download the list of github usernames
* for each user
* create a team with the user name
* add the user to the team
* create a private repository and have the team assigned to it

Get class GitHub Usernames
```{r get_users,eval=FALSE}
users <- read.csv("~/git/gh-usernames/2017_Roster.csv",strip.white = TRUE)
users <- users[users$github.username!="",]
users$team.name <- make.names(paste(users$Last.Name,users$First.Name))
users$Real.Name <- paste(users$First.Name,users$Last.Name)
users
```

Get existing teams on GitHub and compare to imported spreadsheet
```{r get_existing_teams,eval=FALSE}
teams <- gh("/orgs/:org/teams", org="UCDBIS180L-17",.limit=Inf)

current.teams <- data.frame(team.name=sapply(teams, function(x) x$name),
                            team.id=sapply(teams, function(x) x$id))

current.teams

new.teams <- users[! users$team.name %in% current.teams$team.name,]

new.teams
```

Create new teams
```{r new_teams,eval=FALSE}
create.team.out <- apply(new.teams,1,function(team) {
  gh("POST /orgs/:org/teams", 
     org = "UCDBIS180L-17",
     name=team[["team.name"]],
     description=paste("Team for",team[["Real.Name"]],"to turn in BIS180L assignments")
  )
})

new.team.results <- as.data.frame(t(sapply(create.team.out,function(x) c(x$name,x$id))),stringsAsFactors=FALSE)
names(new.team.results) <- c("team.name","team.id")
new.team.results
new.teams <- merge(new.teams,new.team.results,by="team.name")
```

Add members to teams
```{r Add_members,eval=FALSE}
#new.teams <- merge(users,current.teams,by="team.name") #depending on what you are doing you may want to uncoment this
add.members.out <- apply(new.teams,1,function(team) { 
  print(team)
 gh("PUT /teams/:id/memberships/:username",
    id = team[["team.id"]],
     username = team[["github.username"]])
})
#Need to add some checks here
```


Create repository for new or exisiting teams
```{r new_repositories,eval=FALSE}
repo.prefix <- "Assignment_1"

repo.description <- "Repository for BIS180L Assignment 1"

################################################################################
## Choose one of the two below and comment the other out
#teams <- current.teams[!grepl("(Owners|Teachers)",current.teams$team.name),]

teams <- current.teams
#teams <- new.teams[is.numeric(new.teams$team.id),]
#teams <- current.teams[33,]
################################################################################

create.repository.out <- apply(teams,1,function(team) {
  print(team[["team.name"]])
  gh("POST /orgs/:org/repos",
     org="UCDBIS180L-17",
     name=paste(repo.prefix,team[["team.name"]],sep="_"),
     description=paste(repo.description,"from",team[["team.name"]]),
     private="true",
     team_id=team[["team.id"]])
})

```

Give teams push access to their repository...
```{r}
update.repository.out <- apply(teams,1,function(team) {
  print(team[["team.name"]])
  gh("PUT /teams/:id/repos/:org/:repo",
     org="UCDBIS180L-17",
     id=team[["team.id"]],
     repo=paste(repo.prefix,team[["team.name"]],sep="_"),
     permission="push")
})
```

Delete repositories 
```{r new_repositories,eval=FALSE}
repo.prefix <- "Assignment_2"

################################################################################
## Choose one of the two below and comment the other out
#teams <- current.teams[!grepl("(Owners|Teachers)",current.teams$team.name),]
teams <- current.teams
#teams <- new.teams[is.numeric(new.teams$team.id),]
################################################################################

delete.repository.out <- apply(teams,1,function(team) {
  print(team[["team.name"]])
  gh("DELETE /repos/:owner/:repo",
     owner="UCDBIS180L-17",
     repo=paste(repo.prefix,team[["team.name"]],sep="_")
   )
})

```