---
title: Data Structures
layout: post
categories: resources
comments: true
duedate:
related: 2013-03-27-DataTypes
author: Julin Maloof
---

Data Structures
========================================================
This document provides a brief introduciton to *Data Structures*: the different types of objects that R uses for storing data.  See a separate document on *Data Types*

We will cover vectors, matrices, arrays, data frames, and lists.

Vectors
-------
*Vectors* are 1 dimensionsal collections of items of the same data type*.

```{r vector examples}
#construction
a <- 10:20 #a numeric vector
a
#you can add names
names(a) <- letters[1:11]
a
b <- c("julin","siobhan","stacey") # a character vector
b
c <- c(T,T,F) # a logical vector
#testing
#query an object to see if it is a vector:
is.vector(a)
#subetting
#by position
a[4]
a[3:6]
a[c(3,5,7)]
#by name
a["c"]
a[c("b","d","e")]
```
*exception: lists are technically vectors and can hold different data types


Matrices
--------
*Matrices* are 2 dimensional collections of items of the same data type.

```{r matrix examples}
#you can use the matrix() function to make matrices
m <- matrix(1:25,ncol=5)
m
is.matrix(m)
dim(m)
#you can also use rbind() or cbind()
m2 <- rbind(c(1,2,3),
            c(11,12,13),
            c(21,22,23))
m2
```

Arrays
------
*Arrays* are like matrices but have have 2 or more dimensions.  (Matrices are arrays but only have 2 dimensions)
```{r array examples}
#construction
my.array <- array(1:27,dim=c(3,3,3)) # a 3D array
my.array
#testing
is.array(my.array)
#subsetting
#you can access subsets of the array using square brackets
my.array[1,,] # first row, across all columns and the 3rd Dim #note this is now 2 dimensions
my.array[2,2,3] #now 1 dimension
```

Data Frames
-----------
*Data Frames* are 2 dimensional objects, like matrices, **but** each column can hold a different data type.  This is an exceptionaly useful data structure.

```{r data frames}
#construction
df <- data.frame(genotype=c("phyB","Col","phyB","Col"),hyp.length=c(10,3,9,2))
df
#subsetting
#get an individual column with "$"
df$genotype
#or get any subset by []
df[3,2] #second column, third row
df['hyp.length']
#testing
is.data.frame(df)
#converting
as.matrix(df)
as.data.frame(m)

```

Lists
-----
Lists are 1 dimensionsal structures where each item can be any data type.
```{r lists}
#here we create a list that holds many of the earlier objects
my.list <- list(my.matrix=m,my.df=df,my.array=my.array,my.first.vector=a,my.second.vector=b)
my.list
#testing
is.list(my.list)
#subsetting
#you can use the $ extractor
my.list$my.first.vector
#you can use [], which will return a smaller list
my.list[2] # the second item
my.list[2:3] #second and third item.
my.list["my.second.vector"]
#you can use [[]] which will return the item itself
my.list[[2]]
is.data.frame(my.list[[2]])
#compare with
is.data.frame(my.list[2])
#unlist() sometimes is helpful
unlist(my.list[2])
```
*Posted by Julin Maloof*