---
title: More Regular Expressions...
layout: post
categories: exercises
comments: true
duedate: 2013-05-09
related:
author: Mike Covington
---

# Exercise 1: Extracting important info from filenames

You have a dataset and need to extract metadata for IL name, sun/shade treatment, and replicate number from the filenames:

```{r moreregex.01}
filename.1 <- "a IL-9.2.4 crazy SH file rep1 name"
filename.2 <- "anotherIL-12.4sillySUfilerep20name"
```

You'd like to convert the filenames into `IL-9.2.4,SH,rep1` and `IL-12.4,SU,rep20`, which can easily be split on the delimiting comma with `strsplit()`.

Extract this metadata using `sub()` followed by `strsplit` to split the resulting string. You will need to make use of three capture groups.

# Exercise 2: Character conversions

In last week's exercise #2, you were asked to replace all of the "i"s in state names with "y"s. Using `gsub("i", "y", state.name, ignore.case = T)` accomplished that task, but resulted in state names beginning with lower case "y"s.

There is another (non-regex) function that can be used instead: `chartr()` does character translations from one characer to another. To convert all "i"s to "y"s and all "I"s to "Y"s in the state names, we can use the following:

```{r moreregex.02}
data(state)
chartr("Ii", "Yy", state.name)
```

Sidenote: two other non-regex functions used to convert text are `tolower()` and `toupper()`:

```{r moreregex.03}
x <- "Oh my.... Look over there!"
tolower(x)
toupper(x)
```

**Your task**: You want to look cool, so use two of these functions to convert the state names such that all vowells are uppercase and all consonants are lower case.

# Exercise 3: (in progress)



