---
title: Strings and Regular Expressions Answers
layout: post
categories: solutions
comments: true
duedate:
related: 2013-04-21-Text-Searches-Regular-Expressions
author: Julin Maloof
---
String Searching and Manipulation
========================================================
R has extensive tools for searching for text (string) patterns and performing substitutions.  These can be very helpful in manipulation of large data sets.  

These functions can search for exact matches or can make use of an extensive syntax for wildcards, known as *regular expressions*.  I will first describe the search functions and then explain regular expressions.

In this excercise, type all of the code blocks into your computer to see what happens!  Experiment as you go!

Search Functions
----------------

### `grep()`
Use *`grep()`* when you want to search for a single pattern in a vector of strings.
```{r grep example}
#for example: which state names have "i" in them?
data(state)
state.name
#in its basic form, grep returns the position (index) of the items that match the pattern.
grep(pattern="i",x=state.name,ignore.case=T)
#you can also ask grep to return the values that match
grep(pattern="i",x=state.name,ignore.case=T,value=T)
```
The variant *`grepl()`* returns a logical vector of TRUE and FALSE indicating whether or not a match occurred.
```{r grepl example}
grepl(pattern="i",x=state.name,ignore.case=T)
#recalling that TRUE = 1 and FALSE = 0, when can use grepl to quickly 
#determine the number of states that have "i" in them.
sum(grepl(pattern="i",x=state.name,ignore.case=T))
```
**Exercise 1:** Can you figure out how to count the number of states that do not have an "i" in their name (without changing the search pattern)?  There are at least two easy ways.
```{r Ex1}
sum(!grepl(pattern="i",x=state.name,ignore.case=T))
length(grep(pattern="i",x=state.name,ignore.case=T,invert=T))
```


### `match()`
Use *`match()`* when you have two sets of values (lets call them Set A and Set B) and you want to know the (first) positions in Set B that match something that is in Set A.  Match only finds exact matches and can not be used with regular expressions.
```{r match example}
#we have a list of favorite fruits and a list of citrus.  Which of the favorites are citrus?
favorites <- c("peach","bannana","blueberry","orange","plum","strawberry","mandarin")
citrus <- c("kumquat","grapefruit","orange","mandarin","orange","tangerine","tangelo","lemon","lime")
match(favorites,citrus) #what do the numbers returned refer to?
```
Match has an alternative form which I find more convenient: **`%in%`**.  This form returns Trues or Falses indicating whether or not Set A is in Set B.
```{r %in% example}
favorites %in% citrus
#most useful is to use this in combination with square brackets to extract
favorites[favorites %in% citrus]
```
### `sub()`
**`sub()`** performs text substitutions.  Think of it as find and replace.
```{r sub example}
#lets say that we wanted to replace "W" with the phrase "_DoubleU_"
sub("w","_DoubleU_",state.name,ignore.case=T)
```

**Excerise 2:** Try replacing all of the "i"s with "y"s.  Are results what you expected? Why or why not?
```{r Ex2}
sub("i","y",state.name,ignore.case=T)
#Only the first "i" in each word is replaced
```


### `gsub()`
`sub()` only replaces the first occurrence of *pattern*.  **`gsub()`** will replace all occurrences.
```{r gsub example}
gsub("i","y",state.name,ignore.case=T)
#compare to just using sub()
```

Regular Expressions
-------------------
Regular expressions (regexp) are a way of specifying wild cards in the search operations described above. The same syntax (with some modifications) is used in many computer languages including Unix/Linux command line tools, Perl, python, and very helpfully SublimeText2.

### Character classes I
Regexp have codes to match certain _classes_ of characters.

* .  Matches any single character
* \w Matches any character that would be found in a "word" including digits (excludes punctuation and white space)
* \W Is the opposite of \w and matches any non-word character
* \d Matches any digit character
* \D Matches any non-digit character
* \s Matches any white space character
* \S Matches any non space character

The following match specific characters or locations but are worth mentioning here:
* ^  Matches the beginning of a line
* $  Matches the end of a line
* \t tab character
* \n return character

Unless you use additional modifications described below, these match a single character.

How would we find state names that have two "a"s separated by a single additional character?
```{r regex 1}
data(state)
grep("a.a",state.name,value=T,ignore.case=T)
```

**Exercise 3:** Find state names that have a space in their names.
```{r Ex3}
grep(" ",state.name,value=T)
grep("\\s",state.name,value=T)
```

**Exercise 4:** Find state names that begin with "M"
```{r Ex4}
grep("^M",state.name,value=T)
```
### Character classes II
An alternative way of specifying character classes is to enclose them in square brackets.

For example, to find all of the state names that end with a vowel:
```{r regex 2}
grep("[aeiou]$",state.name,value=T)
```
The ^ sign inside of square brackets inverts the search.

**Exercise 5:** Find state names that begin with non-vowels.
```{r Ex5}
grep("^[^aeiou]",state.name,value=T,ignore.case=T)
```
You can specify ranges of characters inside square brackets:

* [0-9] : all digits
* [a-j] : the first 10 (lowercase) letters of the alphabet

There are also a number of predefined character classes.  For example:

[:punct:]
Punctuation characters:
! " # $ % & ' ( ) * + , - . / : ; < = > ? @ [ \ ] ^ _ ` { | } ~.

See `?regex` for additional classes.

Perhaps confusingly these classes must themselves be placed within square brackets
```{r regex 3}
grep("[[:space:]]",state.name,value=T)
```

### Specifying repeats
You can specify the number of times that a character or character class is repeated.

* ?     The preceding item is optional and will be matched at most once.
* *     The preceding item will be matched zero or more times.
* +     The preceding item will be matched one or more times.
* {n}   The preceding item is matched exactly n times.
* {n,}  The preceding item is matched n or more times.
* {n,m} The preceding item is matched at least n times, but not more than m times.

**Exercise 6:** Find the state names that have two consecutive vowels in them.

```{r Ex6}
grep("[aeiou]{2}",state.name,value=T)
```


**Exercise 7:** Using last weeks data set, count the number of babys with the name Stacey or Stacy.  Bonus 1: count Stacey, Stacy, or Staci.  Bonus 2: list the unique set of names that start with "Stac".  The data set can be [downloaded here](http://plyr.had.co.nz/09-user/plyr-tutorial.zip)

```{r Ex7}
bnames <- read.csv("~/Documents/Teaching/RClub/plyr-tutorial/examples/bnames.csv",as.is=T)
#check it first:
unique(grep("Stace?y",bnames$name,value=T))
sum(grepl("Stace?y",bnames$name))
#Bonus1:
unique(grep("Stace?[iy]$",bnames$name,value=T))
sum(grepl("Stace?[iy]$",bnames$name))
#Bonus2:
unique(grep("Stac",bnames$name,value=T))
```


### Alternates
The "|" character can be read as "or" and can be used to specify alternates.
```{r Regex alternates example}
#Example: 
#change the path below to point to "bnames.csv" on your computer
bnames <- read.csv("~/Documents/Teaching/RClub/plyr-tutorial/examples/bnames.csv",as.is=T)
bnames[grep("Stac(y|i)",bnames$name),]
```

**Exercise 8:** Pull out the names "Jonathan" "Jonnie" and "Johnathon" but not other Jon names.  Bonus: Only have Jon listed once in your search string.
```{r Ex8}
unique(grep("^(Johnathon)|(Jon(nie)|(athon))",bnames$name,value=T))

##note that this was harder than intended.  I meant to ask for "Jonathan" "Jonnie" and "Jonathon"
unique(grep("^Jon((nie)|(ath[ao]n))",bnames$name,value=T))
```


### Escapes
What if you want to match a "." or other special character?  The following characters have special meaning in regular expressions: ". \ | ( ) [ { ^ $ * + ?" and if you want to search for them you have to do something special.
```{r regex escapes example}
#say you want all Chrom one ILs
ILs <- c("IL.1.1","IL.2.2","IL.1.3","IL.2.1","IL.11.1","IL.11.3","IL.12.1","IL.12.2")
#the following seems logical at first:
grep("IL.1.",ILs,value=T)
```
What happened?  Remember that "." matches any character.  We need to tell grep that the "." is just a regular character, not a wildcard. To do this we "escape" it by preceding it with a "\".  However, since "\" itself is a special character it must also be escaped. So we use two backslashes. 
```{r regex escapes II}
grep("IL\\.1\\.",ILs,value=T)
#an alternative, if you don't need any regex functionality is to use the argument "fixed=T"
grep("IL.1.",ILs,value=T,fixed=T)
```

### Back references
One of the powerful tools in regexps is the ability to refer back to a previous match.  The item that you want to refer back to is enclosed in parentheses.  You backreference with a backslash and a digit.  A "1" would indicate the first group enclosed in a parentheses, a "2" would refer to the second group, etc.

Suppose we want to find all state names that have letters repeated twice in a row.
```{r}
grep("(.)\\1",state.name,value=T)
#the "." matches any character.  Enclosing that in parentheses designates it as an item that we want to refer back to.  The \\1 refers back to it.
```
**Exercise 9:** Find all state names that have two vowels repeated.
```{r Ex9}
grep("([aeiou])\\1",state.name,value=T,ignore.case=T)
```


I find back references particularly helpful in combination with sub(), because you can use them in your replacement string.

**Exercise 10:** For all two-worded state names reverse the order of the two words and add a comma between the words.  ("North Carolina" should become "Carolina, North")
```{r Ex10}
sub("([[:alpha:]]+) ([[:alpha:]]+)","\\2, \\1",state.name)
```


**Exercise 11:** For all two-worded state names, abbreviate the first word to be the first letter, followed by a "." ("North Carolina" becomes "N. Carolina").
```{r Ex11}
sub("^(\\w)\\w+\\s","\\1. ",state.name)
```
