---
title: Homework--Data Types and Reshape
layout: post
categories: solutions
comments: true
duedate: 2013-04-04
author: Julin Maloof
---

R Homework for April 4, 2013
========================================================

## Section 1: Data Types
Questions 1 to 6 refer to the file [ELF3s\_bolting\_Rclub.csv]({{ site.baseurl }}/downloads/ELF3s_bolting_Rclub.csv). Import it now.

### Q1
What data types are represented in each column?
```{r 2013-04-03-Data-Solutions-Q1}
data <- read.csv("~/Documents/Teaching/RClub/webSite/downloads/ELF3s_bolting_Rclub.csv")

#one way is to do a summary and figure it out from the way the output is formatted
summary(data) 
#factor: Plant, diameter, treatment, genotype, transformation, col,
#numeric: n_leaves, flowering_time, flat, row

#alternative:
str(data)



```



### Q2
a) Are there any columns that you think have the wrong data type?

*Yes*

b) Which ones? *see below*

c) Why?

*Diameter should be numeric because we would want to analyze the actual measurements.  Flat and row should be factor becuase these are used to group plants.*

### Q3
How would you change the columns to their correct types?

```{r 2013-04-03-Data-Solutions-Q3}
#For row and flat:
data$flat <- as.factor(data$flat)
data$row <- as.factor(data$row)

#It isn't so simple for diameter.
#First is to determine why diameter came in as a factor instead of numeric.
#There must be a non-numeric value somewhere in there and we need to find it.

#can we just use as.numeric?
as.numeric(data$diameter)
#No! this returns the integer representation of the factor levels.

#instead:
as.numeric(as.character(data$diameter))

#or:
as.numeric(levels(data$diameter)[data$diameter])

#The warning tells us that there was a problem in the conversion, 
#this is probably our problem observation

#find it with:
data[is.na(as.numeric(as.character(data$diameter))),] 
#Aha an extra space!

#an alternative way to find this would be:
grep("[^0-9\\.]+",data$diameter,value=T)
#which translates into find any item that contains something other than the digits 0-9 and a decimal point.
#We will have a session on grep at some later point.

#fix it with:
data$diameter <- as.character(data$diameter)
data$diameter[data$diameter=="43. 68"] <- 43.68
data$diameter <- as.numeric(data$diameter)

#or, more generally:
data$diameter <- as.numeric(sub(" ","",data$diameter))
#this searches for a space and substitutes an empty string
#We don't have to convert to a character because sub does this automatically

#An alternative way to change data types in the data frame, from Aashish
data_mod <- transform(data,
                      diameter = as.numeric(as.character(diameter)),
                      flat = as.factor(flat),
                      row = as.factor(row))

```

### Q4
Are there any obvious mistakes in this data frame beyond what you might have found in answering Q3?
```{r 2013-04-03-Data-Solutions-Q4}

#quick check is with summary
summary(data)
#Plants with diameter 0 and 0 leaves; seems unusual
#Also there are two entries for Sh2.H4.1h

#lets look at the duplicated entry a little more closely
data[duplicated(data$Plant),] #only shows us the second row
dup.plant <- data$Plant[duplicated(data$Plant)]
data[data$Plant==dup.plant,]

#since the values are the same we can just remove it with:
data <- data[!duplicated(data$Plant),]
# OR
data <- unique(data)

#what about the strange plants with zero diameter?
#might be worth doing a quick visual check of outliers:
library(reshape)
library(ggplot2)
melt.data <- melt(data,measure.vars=c("diameter","n_leaves","flowering_time"))
pl <- ggplot(melt.data,aes(x=variable,y=value))
pl + geom_jitter() + facet_wrap( ~ variable,scales="free_y")

#flowering time and n_leaves of 0 seems odd.  n_leaves of 50 perhaps a bit strange
#lets look at these rows
data[data$diameter==0,]
#remove row with 0 measurements
data <- data[data$diameter>0,]

#look at row with more than 50 leaces
data[data$n_leaves>50,] 
#also has among the largest diameters and latest flowering times, so seems real.
```


### Q5
Make the "Sha" genotype the reference level for the genotype column
```{r 2013-04-03-Data-Solutions-Q5}
data$genotype <- relevel(data$genotype,ref="Sha")
```


### Q6
Change the order of levels in "trasformation" to be Sh1, Sh2, Sh3, Ba1, Ba3
```{r 2013-04-03-Data-Solutions-Q6}
data$genotype <- factor(data$genotype,levels=c("Sh1", "Sh2", "Sh3", "Ba1", "Ba3"))
```


## Section 2: Reshape
The remaining questions deal with the *reshape* package and the tomato dataset.

* Read my post on reshape [on the Rclub website](http://mfcovington.github.com/r_club/resources/2013/03/28/Reshape/)
* Additional information on reshape is in section 9.2 of the ggplot book
* import the standard tomato data set.

```{r}
tomato <- read.csv("~/Documents/Teaching/RClub/TomatoRClub.csv")
```

### Q7
What are the id variables and measure variables in the Tomato data set?
```{r 2013-04-03-Data-Solutions-Q7}
head(tomato)
#id.vars: shelf, flat, col, row, acs, trt, ndvi, lat, lon, alt, species, who
#measure vars: hyp, int1, int2, int3, int4, intleng, totleng, petleng, leafleng, leafwid, leafnum
```


### Q8
Subset the tomato data set to keep the int1-int4 measurements and the relevant metadata.
```{r 2013-04-03-Data-Solutions-Q8}
tom.small <- tomato[,c(1:6,10:13,20:25)]
head(tom.small)
```


### Q9
Without melting or casting your new data frame, calculate the mean of each internode. *Hint: use `apply()`*
```{r 2013-04-03-Data-Solutions-Q9}
apply(tom.small[,7:10],2,mean,na.rm=T)

#OR

colMeans(tom.small[,7:10],na.rm=T) #Thanks Upendra
```


### Q10
Melt the new data frame.
```{r 2013-04-03-Data-Solutions-Q10}
tom.melt <- melt(tom.small,measure.vars=7:10,variable_name="internode")
summary(tom.melt)
```

### Q11
Use cast to obtain the mean for each internode.
Q11
```{r 2013-04-03-Data-Solutions-Q11}
cast(tom.melt, ~ internode, mean, na.rm=T)
```

### Q12
Use cast to obtain the mean for each internode for each species.
```{r 2013-04-03-Data-Solutions-Q12}
cast(tom.melt,  internode ~ species, mean, na.rm=T)
```

### Q13
Use cast to obtain the mean for each internode for each species under each treatment.
```{r Q13}
cast(tom.melt,  internode ~ species ~ trt, mean, na.rm=T)
```

### Q14
Create a boxplot for each combination of species, internode, and treatment.
```{r 2013-04-03-Data-Solutions-Q14}
pl <- ggplot(tom.melt,aes(x=internode,y=value,color=trt))
pl <- pl + geom_boxplot()
pl + facet_grid(species ~ .,scales="free_y")
```

##Bonus: unmelt your data with cast
If you include an indexing variable in your original data frame, you can use case to unmelt it.
```{r 2013-04-03-Data-Solutions-unmelt}
tom.small$index <- 1:nrow(tom.small)
tom.melt2 <- melt(tom.small,measure.vars=7:10)
head(tom.melt2)
tom.small2 <- cast(tom.melt2)
tom.small2 <- tom.small2[order(tom.small2$index),]
dim(tom.small)
dim(tom.small2)
head(tom.small)
head(tom.small2)
```
*Posted by Julin Maloof*
