---
title: Data Types
layout: post
categories: resources
comments: true
duedate:
---

Data Types
========================================================
This document provides a brief introduction to *Data Types*: how data is represented in R.  A separate document discusses *Data Structures*: the kinds of objects that data can be stored in.

R categorizes data as different types: numeric, character, factor, boolean/logical, date, etc.

Numeric
-------
The *numeric* data type is for numbers.  Numbers can be used for calculations.  Some relevant functions are:
```{r numeric}
my.numbers <- c(4,5) # a numeric vector
sum(my.numbers)
is.numeric(my.numbers)
my.text.numbers <- c("4","5") # a character vector
sum(my.text.numbers)
is.numeric(my.text.numbers)
as.numeric(my.text.numbers)
sum(as.numeric(my.text.numbers))
```

Character
---------
The *character* data type is for representing text.  You can tell if something is a charcter type in R because it will be enclosed in quotes.
```{r character}
my.text <- c("Julin","Maloof")
my.text
is.character(my.text)
#converting numbers to characters
my.numbers
as.character(my.numbers) # note the quotation marks
```

Factor
------
The *factor* data type is used to designate groups.  Factors have associated levels that represent each possible group available in the factor.
```{r factors}
genotypes <- factor(c("wildtype","mutant1","mutant2","wildtype"))
genotypes #note that the levels are listed and the values are not in quotes
class(genotypes)
is.factor(genotypes)
is.character(genotypes)
as.character(genotypes) # note the quotes
levels(genotypes) #alpabetical by default
nlevels(genotypes)
#often you want wildtype to be the first level
genotypes <- relevel(genotypes,ref="wildtype")
levels(genotypes)
#or maybe you want to have a custom order for everything
genotypes <- factor(genotypes,levels=c("mutant2","wildtype","mutant1"))
genotypes
```

Logical
-------
The *logical* data type is used for true/false values
```{r logical}
my.boolean <- c(F,T,T,F) #you could also use TRUE and FALSE
my.boolean
is.logical(my.boolean)
# The '!' reverses the values (logical NOT)
!my.boolean
#logicals can be used in extraction
genotypes[my.boolean]
genotypes[genotypes=="wildtype"] #here you are creating a logical inside the square brackets.
genotypes=="wildtype"
#conversions:
as.numeric(my.boolean) # 1 is true, 0 is false.  This can be useful for summing, ie
sum(genotypes=="wildtype")
as.logical(c(1, 0, 1, 0))
as.character(my.boolean)
#converting from text
as.logical(c("T","True","true","TRUE","F","False","FALSE","false","not logical"))
```

*Posted by Julin Maloof*